Nodal Discontinuous Galerkin Method for the Euler Equations of Fluid Dynamics

### 2017-06-30 Documents/Euler/dgEulerCurvilinear.tex
Some text on positivity limiter

@@

For general curvilinear spatial coordinates, the non-relativistic Euler equations for gas dynamics are \citep[][see, e.g., \citet{} and references therein for a fully general relativistic treatment]{leveque_2002}
\begin{equation}
  \pderiv{\rho}{t}
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\rho\,v^{i}\,\Big)=0,
  \label{eq:massConservation}
\end{equation}
\begin{equation}
  \pderiv{S_{j}}{t}
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,P^{i}_{~j}\,\Big)
  =\f{1}{2}\,P^{ik}\,\pderiv{\gamma_{ik}}{x^{j}}-\rho\,\pderiv{\Phi}{x^{j}},
  \label{eq:momentumConservation}
\end{equation}
\begin{equation}
  \pderiv{E}{t}
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\big(E+p\big)\,v^{i}\,\Big)
  =-S^{i}\,\pderiv{\Phi}{x^{i}},
  \label{eq:energyConservation}
\end{equation}
\begin{equation}
  \pderiv{n_{e}}{t}
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,n_{e}\,v^{i}\,\Big)=0,
  \label{eq:electronConservation}
\end{equation}
where $\rho$ is the (baryon) mass density, $S^{i}=\rho\,v^{i}$ is the momentum density, $P^{ij}=\rho\,v^{i}\,v^{j}+p\,\gamma^{ij}$ the stress tensor, $E=e+\f{1}{2}\,\rho\,v_{i}\,v^{i}$ the total (internal plus kinetic) energy density, and $n_{e}$ the electron density.  
The pressure is related to the density, internal energy density, and electron density by an equation of state (EoS); i.e., $p=p(\rho,e,n_{e})$.  
(We also define the specific internal energy by $\epsilon=e/\rho$.)
The Newtonian gravitation potential, $\Phi$, is obtained from solving
\begin{equation}
  \f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\gamma^{ij}\pderiv{\Phi}{x^{j}}\,\Big)=4\pi\,\rho.  
  \label{eq:poisson}
\end{equation}

@@

>>>>

\clearpage

@@

For general curvilinear spatial coordinates, the non-relativistic Euler equations for gas dynamics are %\citep[][see, e.g., \citet{} and references therein for a fully general relativistic treatment]{leveque_2002}
\begin{equation}
  \pderiv{\rho}{t}
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\rho\,v^{i}\,\Big)=0,
  \label{eq:massConservation}
\end{equation}
\begin{equation}
  \pderiv{S_{j}}{t}
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,P^{i}_{~j}\,\Big)
  =\f{1}{2}\,P^{ik}\,\pderiv{\gamma_{ik}}{x^{j}}-\rho\,\pderiv{\Phi}{x^{j}},
  \label{eq:momentumConservation}
\end{equation}
\begin{equation}
  \pderiv{E}{t}
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\big(E+p\big)\,v^{i}\,\Big)
  =-S^{i}\,\pderiv{\Phi}{x^{i}},
  \label{eq:energyConservation}
\end{equation}
\begin{equation}
  \pderiv{n_{e}}{t}
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,n_{e}\,v^{i}\,\Big)=0,
  \label{eq:electronConservation}
\end{equation}
where $\rho$ is the (baryon) mass density, $S^{i}=\rho\,v^{i}$ is the momentum density, $P^{ij}=\rho\,v^{i}\,v^{j}+p\,\gamma^{ij}$ the stress tensor, $E=e+\f{1}{2}\,\rho\,v_{i}\,v^{i}$ the total (internal plus kinetic) energy density, and $n_{e}$ the electron density.  
The pressure is related to the density, internal energy density, and electron density by an equation of state (EoS); i.e., $p=p(\rho,e,n_{e})$.  
(We also define the specific internal energy by $\epsilon=e/\rho$.)
The Newtonian gravitation potential, $\Phi$, is obtained from solving
\begin{equation}
  \f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\gamma^{ij}\pderiv{\Phi}{x^{j}}\,\Big)=4\pi\,\rho.  
  \label{eq:poisson}
\end{equation}

@@

In the case of the nuclear EoS, the thermodynamic quantities (e.g, pressure) are tabulated in terms of mass density $\rho$, temperature $T$, and electron fraction $Y_{e}$.  
Obviously, the tables are of finite extent in these independent variables; i.e., $\rho\in[\rho_{0},\rho_{1}]$, $T\in[T_{0},T_{1}]$, and $Y_{e}\in[Y_{e,0},Y_{e,1}]$.  
During stellar core-collapse, the internal energy density has a dominant contribution from degenerate electrons, which is independent of the temperature.  
During core bounce and shock formation, we have found that the numerical scheme (e.g., during reconstruction of element interfaces) can lead to internal energy densities that fall below the lower boundary of the tabulated data and failure to obtain a temperature, which is needed to evaluate the pressure.  
(Extending the table to lower values of temperature does not help since the internal energy density is practically independent of the temperature.)

For the solution to the Euler equations, we define the set of physically admissible states as
\begin{equation}
  \cG=\big\{\,\bU~|~\rho\ge\rho_{0}, e\ge e_{0}(\rho,T_{0},n_{e}),n_{e}\ge n_{e,0}(\rho,Y_{e,0})\,\big\}
\end{equation}
